---
title: "Police Complaints in Chicago"
subtitle: "Presentation subtitle (if any)"
author: "Elly Beckerman, Sokona Mangane"
institute: "Bates College"
date: "`r Sys.Date()`"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    nature:
      ratio: "16:9"
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r load-packages, include = FALSE}
# Add any additional packages you need to this chunk
library(tidyverse)
library(tidymodels)
library(palmerpenguins)
library(knitr)
library(xaringanthemer)
```

```{r setup, include=FALSE}
# For better figure resolution
knitr::opts_chunk$set(fig.retina = 3, dpi = 300, fig.width = 6, fig.asp = 0.618, out.width = "75%", warning = FALSE, echo = FALSE)
```

```{r load our packages, include = FALSE}
library(knitr)
library(tidyverse)
library(readr)
library(tidyverse)
library(forcats)
library(kableExtra)
library(sf)
library(leaflet)
#install.packages("naniar")
library(naniar)
# set default options

```

```{r Reading in the data, echo=FALSE, message=FALSE, warning = FALSE}

district_demographics <- read_csv("https://raw.githubusercontent.com/invinst/chicago-police-data/master/data/context_data/district_demographics/Districts.csv")

active_duty_officers <- read_csv("https://raw.githubusercontent.com/invinst/chicago-police-data/master/data/context_data/CPD%20Employees%20active-duty-only.csv")

#The officers accused
complaints_accused <- read_csv("../data/complaints-accused_2000-2016_2016-11.csv")

#those who made the complaints
complainants <- read_csv("../data/complaints-complainants_2000-2016_2016-11 (1).csv")

#What complaints were made (more categorical data)
complaints_made <- read_csv("../data/complaints-complaints_2000-2016_2016-11 (1).csv")

units <- read.csv("https://raw.githubusercontent.com/invinst/chicago-police-data/master/data/context_data/current_and_past_units/Current_and_Past_CPD_Units_2016-05-06.csv")

```

```{r background-image, include = FALSE}
style_xaringan(
  title_slide_background_image = "img/web-chicago-skyline.webp"
)
```

# Our Research Questions
- What are the racial makeups of different neighborhoods in Chicago?

--

- In which neighborhoods are there the most police complaints per capita?

--

- What are the outcomes of police complaints?

---

class: inverse, center, middle

# Section title

---

# Introducing our data

- Data collected by the Invisible Institute, an investigative journalism company

--

- The Invisible Institute sued the city of Chicago for the release of complaint records against the Chicago Police Department

--

- The files were released in 2016, revealing over 100,000 complaint records to the public

---

# Complaints per capita
.pull-left[
- Some text
- goes here
]
.pull-right[
```{r data-joins, fig.alt = "Scatterplot of police complaints per capita by district name colored by racial majority. Visualization shows", echo = FALSE}
district_complaints <- complaints_accused %>%
  filter(current_unit %in% 1:25) %>%
  group_by(current_unit) %>%
  summarise(n = n()) %>%
  arrange(desc(n))

total_district_complaints <- full_join(district_complaints,
                                       district_demographics,
                                       by = c("current_unit" = "District_No")) %>%
  mutate(complaints_per_capita = n/Population)

total_district_complaints %>%
  filter(is.na(District_Name) == FALSE) %>%
         ggplot(mapping = aes(
                  x = fct_reorder(District_Name, complaints_per_capita),
                  y = complaints_per_capita, 
                color = Majority)) +
  geom_point() +
  geom_segment(aes(x = fct_reorder(District_Name, complaints_per_capita), 
                   xend = fct_reorder(District_Name, complaints_per_capita), 
                   y = 0, yend = complaints_per_capita)) +
  coord_flip() +
  labs(title = "Complaints per Capita by District Name",
       subtitle = "Colored by Racial Majority",
       x = "District Name",
       y = "Complaints Per Capita")
```
  ]
---

# Proportion final findings
.pull-left[
- Some text
- goes here
]
.pull-right[
```{r proportion-findings-vis, echo = FALSE}
district_complaints_1 <- complaints_accused %>%
  filter(current_unit %in% 1:25) %>%
  group_by(current_unit)

total_district_complaints_findings <- full_join(district_complaints_1,
                                       district_demographics,
                                       by = c("current_unit" = "District_No"))

#stat = "identity"

data1 <- total_district_complaints_findings %>%
  mutate(final_decision = 
           as.factor(case_when(final_finding %in% "SU" ~ "Sustained",
                               final_finding %in% "DIS" ~ "Sustained",
                               is.na(final_finding) == TRUE ~ "Missing",
                               final_finding %in% "NAF" ~ "No Affidavit or Cooperation",
                               final_finding %in% "NC" ~ "No Affidavit or Cooperation",
                               final_finding %in% "NS" ~ "Not Sustained",
                               final_finding %in% "EX" ~ "Not Sustained",
                               final_finding %in% "UN" ~ "Not Sustained",
                               ))) %>%
  mutate(District_Name = fct_relevel(District_Name,
                                     "Central",
                                     "Englewood",
                                     "Harrison",
                                     "Calumet",
                                     "Grand Crossing",
                                     "Gresham",
                                     "Austin",
                                     "Morgan Park",
                                     "Wentworth",
                                     "Near North",
                                     "South Chicago",
                                     "Ogden",
                                     "Lincoln",
                                     "Near West",
                                     "Chicago Lawn",
                                     "Deering",
                                     "Shakespeare",
                                     "Town Hall",
                                     "Rogers Park",
                                     "Grand Central",
                                     "Albany Park",
                                     "Jefferson Park"
                                     )) %>%
  group_by(final_decision, District_Name)  %>%
  summarize(n = n())

data1 %>%
  group_by(final_decision) %>%
  summarize(n = n())

#reorder so that missing is at the end and change colors, take out the NA
  data1 %>%   
  filter(is.na(District_Name) == FALSE) %>%
  ggplot(aes(fill = factor(final_decision, levels = c("Missing", "Not Sustained", "No Affidavit or Cooperation", "Sustained")),
                           x = fct_relevel(District_Name), district_levels,
                           y = n)) +
  geom_bar(position = "fill", stat = "identity") +
  theme_minimal() +
  scale_fill_viridis_d() +
  coord_flip() +
   labs(title = "Proportion of Final Findings",
        subtitle = "By District",
        x = "District Name",
        y = "Proportion of Complaints",
        fill = "Final Decision"
        ) 

```
]
---

# Sustained complaints
.pull-left[
- Some text
- goes here
]
.pull-right[
```{r complaints for sustained decisions, echo = FALSE, warning = FALSE}

sustained_data <- total_district_complaints_findings %>%
  group_by(final_finding, District_Name, Majority, Population) %>%
  filter(final_finding == "SU") %>%
  filter(is.na(District_Name) == FALSE) %>%

  summarize(n = n()) %>%
  mutate(complaints_per_capita = n/Population) 
  

  ggplot(data = sustained_data, 
                mapping = aes(
                  x = fct_reorder(District_Name, complaints_per_capita),
                  y = complaints_per_capita, 
                color = Majority)) +
  geom_point() +
    geom_segment(aes(x = fct_reorder(District_Name, complaints_per_capita), 
                     xend = fct_reorder(District_Name, complaints_per_capita), 
                   y = 0,
                   yend = complaints_per_capita)) +
  coord_flip() +
  labs(title = " Sustained Complaints per Capita by District Name",
       subtitle = "Colored by Racial Majority",
       x = "District Name",
       y = " Sustained Complaints Per Capita")
```
]

---
# Layouts

You can use plain text

- or bullet points

.pull-left[
or text in two columns $^*$
]
.pull-right[
- like
- this
]

.footnote[
[*] And add footnotes
]

---

# Code

```{r boring-regression}
# a boring regression
model <- lm(dist ~ speed, data = cars)
tidy(model)
glance(model)
```

---

# Plots

```{r recode-species, echo = FALSE}
# In this chunk I'm doing a bunch of analysis that I don't want to present 
# in my slides. But I need the resulting data frame for a plot I want to present.
iris_modified <- iris %>%
  mutate(Species = fct_other(Species, keep = "setosa"))
```

```{r plot-iris, echo = FALSE}
# Code hidden with echo = FALSE
# Uses modified iris dataset from previous chunk
# Play around with height and width until you're happy with the look
ggplot(data = iris_modified, mapping = aes(x = Sepal.Width, y = Sepal.Length, color = Species)) +
  geom_point() + 
  theme_minimal() # theme options: https://ggplot2.tidyverse.org/reference/ggtheme.html
```

---

## Plot and text

.pull-left[
- Some text
- goes here
]
.pull-right[
```{r warning=FALSE, out.width="100%", fig.width=4, echo=FALSE}
# see how I changed out.width and fig.width from defaults
# to make the figure bigger
ggplot(penguins, aes(x = bill_length_mm, y = species, color = species)) +
  geom_boxplot() +
  theme_minimal()
```
]

---

# Tables

If you want to generate a table, make sure it is in the HTML format (instead of Markdown or other formats), e.g.,

```{r iris-table, echo = FALSE}
kable(head(iris), format = "html")
```

---

# Images

```{r castle, echo = FALSE, out.width = "60%", fig.align = "center", fig.cap = "Image credit: Photo by Jörg Angeli on Unsplash."}
include_graphics("https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80")
```

Or you can also include a full page image. See next slide.

---

background-image: url(https://images.unsplash.com/photo-1535448033526-c0e85c9e6968?ixlib=rb-1.2.1&ixid=eyJhcHBfaWQiOjEyMDd9&auto=format&fit=crop&w=1650&q=80)

---

# Math Expressions

You can write LaTeX math expressions inside a pair of dollar signs, e.g. &#36;\alpha+\beta$ renders $\alpha+\beta$. You can use the display style with double dollar signs:

```
$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$
```

$$\bar{X}=\frac{1}{n}\sum_{i=1}^nX_i$$

Limitations:

1. The source code of a LaTeX math expression must be in one line, unless it is inside a pair of double dollar signs, in which case the starting `$$` must appear in the very beginning of a line, followed immediately by a non-space character, and the ending `$$` must be at the end of a line, led by a non-space character;

1. There should not be spaces after the opening `$` or before the closing `$`.

1. Math does not work on the title slide (see [#61](https://github.com/yihui/xaringan/issues/61) for a workaround).

---

# Feeling adventurous?

- Want to find out more about `xaringan`? See https://slides.yihui.name/xaringan/#1.

- You are welcomed to use the default styling of the slides. In fact, that's what I expect majority of you will do. You will differentiate yourself with the content of your presentation.

- But some of you might want to play around with slide styling. The 
`xaringanthemer` provides some solutions for this that: https://pkg.garrickadenbuie.com/xaringanthemer.

- And if you want more bells and whistles, there is also `xaringanExtra`: https://pkg.garrickadenbuie.com/xaringanExtra.
